---
title: "Data exploration Europe"
author: "Nik Seldon"
date: "26 3 2020"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Phase of exponential growth in Europe

#### Include R libraries
```{r message=FALSE, warning=FALSE,echo=TRUE}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(lubridate)
```


```{r message=FALSE, warning=FALSE,echo=FALSE}
if(.Platform$OS.type == "windows") {
     setwd("C:/Users/domin/OneDrive/covid/")} else {
         setwd("~/OneDrive/covid/")}

cwd<-getwd()
```

#### Getting Data

COVID-19 ata provided by the Johns Hopkins University Center for Systems Science and Engineering (JHU CSSE), 
and is taken from their Github repository[https://github.com/CSSEGISandData/COVID-19] https://github.com/CSSEGISandData/COVID-19), Novel Coronavirus (COVID-19) Cases.

```{r message=FALSE, warning=FALSE,echo=TRUE}
#URLS
github<-"https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/"
confirmedURL_JHU <-paste(github,"csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv?accessType=DOWNLOAD", sep="")
deathURL_JHU<-paste(github,"csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv?accessType=DOWNLOAD", sep="")
```

```{r message=FALSE, warning=FALSE}
#DOWNLOAD
dateDownloaded<-Sys.Date()%>%as.Date("%d-%m-%y")
pathConfirmed<-paste(cwd,"/data/confirmedCases_",dateDownloaded,".csv",sep="")
pathDeath<-paste(cwd,"/data/deathsCases_",dateDownloaded,".csv",sep="")

if(.Platform$OS.type == "windows") {
  download.file(confirmedURL_JHU,destfile = pathConfirmed)
  download.file(deathURL_JHU,destfile = pathDeath)
  #download.file(gitGermanyURL,destfile = pathGithubGermany)
  } else { 
    download.file(confirmedURL_JHU,destfile = pathConfirmed, method = "curl")
    download.file(deathURL_JHU,destfile = pathDeath, method = "curl")
    #download.file(gitGermanyURL,destfile = pathGithubGermany, method = "curl")
    }


#download.file(confirmedURL_JHU,destfile = pathConfirmed, method = "curl")
#download.file(deathURL_JHU,destfile = pathDeath)
```
Here is the code used to aquire the data from JHU CSSE Github-repository

```{r message=FALSE, warning=FALSE}
#Read Saved File
confirmedCases_JHU <- read_csv(pathConfirmed)%>%
    as_tibble()%>%
    rename(province = "Province/State", country_region = "Country/Region")%>% 
    pivot_longer(-c(province, country_region, Lat, Long), names_to = "Date", values_to = "cumulative_cases") %>% 
    mutate(Date=as.Date(mdy(Date)))%>%
    select(-c( Lat, Long)) %>%
    arrange(country_region)

confirmedCasesTmp<-as_tibble()    
countries<-list("Italy", "Germany", "Switzerland", "Spain", "France", "Netherlands")
confirmedCasesSelected<-as_tibble()

for(i in seq_along(countries)){
  confirmedCasesTmp<-confirmedCases_JHU%>%
    filter(country_region == countries[i], province==countries[i] | is.na(province))%>%
    mutate(incident_cases = c(0, diff(cumulative_cases)))
  confirmedCasesSelected<-rbind(confirmedCasesSelected,confirmedCasesTmp)
}
```

### Phase of Exponential Growth

Currently, all countries in Europe are in the epidemic phase of were the increase of incidence can be described by a law of exponential growth. epidemiologist may write <em>I<sub>d</sub>=I<sub>0</sub> * C<sup>d</sup></em>, where <em>I<sub>d</sub>, I<sub>0</sub></em> are the number of infected people at day <em>d</em> and the initial value of infected people. The factor <em>C</em> contains the average number <em>E</em> of people, an infected person is exposed to each day, and the propability <em>p</em> of each exposure becoming an infection. The exponential growth can be slowed down, when <em>E</em> and/or <em>p</em> can be minimised. Instead of <em>C</em>, Physicists would use the exponential base <em>e</em>: <em>I<sub>d</sub>=I<sub>0</sub> * e<sup>(k\*d)</sup></em>. The logarithmic plot of cumulative incidences <em>ln(I<sub>d</sub>/I<sub>0</sub>)=k*d</em> over days yield a straight line, with the slope <em>k</em>. Finally the doubling time - the time it reqiures the number of infected has doubled -  can be calculated: <em>T<sub>2</sub> = ln2/k</em>. In the following figure, the model fit is shown as a red line. 

```{r warning=FALSE}
title<-paste("COVID-19 daily incident  confirmed cases\n data downloaded from Johns Hopkins University: ", dateDownloaded)
caption<-"Source: Johns Hopkins CSSE Novel coronavirus COVID-19 (2019-nCoV)\n data repository: https://github.com/CSSEGISandData/COVID-19"

ggplot(confirmedCasesSelected, aes(x = Date, y = cumulative_cases))+ 
  geom_line() + 
  geom_point()+
  geom_smooth(data=filter(confirmedCasesSelected, Date >= "2020-03-01", Date <= "2020-03-15"), method = "lm",formula =y ~x, col="blue",linetype="dashed")+
  geom_smooth(data=filter(confirmedCasesSelected, Date >= "2020-03-15",Date <= "2020-03-26"), method = "lm",formula =y ~x, col="red",linetype="dashed")+
  facet_wrap(country_region ~ .)+
  labs(y = "Cumulative confirmed cases", title = title, 
       subtitle = "", caption = caption) +
  xlim(c(as.Date('2020-02-15 00:00:00', format = "%Y-%m-%d %H:%M:%S"),
         as.Date('2020-04-01 00:00:00', format = "%Y-%m-%d %H:%M:%S")))+
  scale_y_log10(limits = c(1,1e5))+
  theme(legend.position = "none", legend.title = element_blank())

```





